Supporting information

### 2019-08-26 EAshiklomanovSA2.Rmd
Respond to sampling bias comment

```{r}
boot_data <- bootstrap_missing_hier(result, Y, grp_m)
boot_means <- apply(boot_data, c(1, 2), mean)
boot_lo <- apply(boot_data, c(1, 2), quantile, probs = 0.025)
boot_hi <- apply(boot_data, c(1, 2), quantile, probs = 0.975)
boot_means[!is.na(Y)] <- NA
boot_lo[!is.na(Y)] <- NA
boot_hi[!is.na(Y)] <- NA

@@

>>>>

```{r}
boot_data <- bootstrap_missing_hier(result, Y, grp_m, progress = FALSE)
boot_means <- apply(boot_data, c(1, 2), mean)
boot_lo <- apply(boot_data, c(1, 2), quantile, probs = 0.025)
boot_hi <- apply(boot_data, c(1, 2), quantile, probs = 0.975)
boot_means[!is.na(Y)] <- NA
boot_lo[!is.na(Y)] <- NA
boot_hi[!is.na(Y)] <- NA

@@

# Method S2: Biased mean estimates in non-representatively sampled bivariate data

One important result of this manuscript is that accounting for trait covariance can reduce biases in trait sampling.
We can demonstrate this with an idealized case where we know the correct answer.
Consider two traits "A" and "B", with mean zero, unit variance, and strong covariance ($R = 0.8$):

```{r}
library(mvtraits)
mu <- c(0, 0)
sigma <- matrix(c(1, 0.8, 0.8, 1), nrow = 2, ncol = 2)
n <- 5000
sim <- random_mvnorm(n, mu, sigma)
colnames(sim) <- c("A", "B")
plot(sim[, "A"], sim[, "B"], xlab = "A", ylab = "B", pch = 19, cex = 0.3)
```

Now, suppose that trait B is sampled with a strong preference for low values.
We can simulate this by removing a random sample of B weighted by its value along the normal distribution cumulative distribution function.

```{r}
sim2 <- sim
sim2[sample(n, n / 2, prob = pnorm(sim2[, "B"])), "B"] <- NA

colMeans(sim2, na.rm = TRUE)

plot(density(sim[, "B"]), col = "black",
     main = "Kernel density of trait B",
     xlab = "Trait B")
lines(density(sim2[, "B"], na.rm = TRUE), col = "red")
abline(v = mean(sim[, "B"]), col = "black", lty = "dashed")
abline(v = mean(sim2[, "B"], na.rm = TRUE), col = "red", lty = "dashed")
legend("topright", c("True", "Observed"), lty = 1, col = c("black", "red"))
```

We can see that the univariate estimate of trait B, -0.43, is systematically biased.
Now, let's fit these data using our multivariate model.

```{r}
fit <- fit_mvnorm(sim2)
subset(fit$summary_table,
       variable == "mu",
       select = c("variable", "index", "Mean", "2.5%", "97.5%"))
```

We can see that the mean estimate for B is still negatively biased, but much less so than in the univariate case.
